Laplacian envelope

### 2022-09-05 Notes/laplace.tex
eigenmoodes for uni serial

% \DeclareFontSubstitution{LS2}{stix}{m}{n}
% \DeclareSymbolFont{xlargesymbols}{LS2}{stixex}{m}{n}
% \DeclareMathDelimiter{\lAngle}{\mathopen}{xlargesymbols}{"EC}{xlargesymbols}{"12}
% \DeclareMathDelimiter{\rAngle}{\mathclose}{xlargesymbols}{"ED}{largesymbols}{"13}
% \DeclareMathDelimiter{\lVvert}{\mathopen}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
% \DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}

\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%probability distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diagonal first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%KKT multipliers
\newcommand{\kktm}{\mu}
\newcommand{\kkt}{\boldsymbol{\kktm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\cnst}{\gamma}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

@@

So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
Because we think that \cref{eq:sqrtbound} can be improved, we will rewrite it as
%
\begin{equation}\label{eq:gammabnd}
  \CI_a \sqrt{r\tau_a} \leq \sqrt{\cnst}.
\end{equation}
% 
We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is less than 2.
Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
Later, looking at uniform serial models in \cref{sec:heuristicenv} we will see near \cref{eq:henvapprox} that \(\gamma\) is at least 1.18.

@@

The first of these equations is solved by
%
\begin{equation}\label{eq:detagensol}
  \delta\etwm_i = B\e^{\beta(i-1)} + C\e^{\beta(1-i)},
  \qquad\text{where}\qquad
  s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}}
    = \cosh (\beta) - 1,
\end{equation}
%
with \(B\) and \(C\) determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
%
\begin{equation}\label{eq:unicoeffs}
  B = -\frac{1}{s(1 + \e^{-\beta}) \cosh \prn{m \beta}},
  \qquad
  C = -\frac{1}{s(\e^{\beta} + 1) \cosh \prn{m \beta}}.
\end{equation}
%
This can be used to compute the Laplace transform
%
\begin{equation}\label{eq:unilaplace}
  A(s) = \frac{\etwm_1(s)}{m}
   = \frac{1}{m}\prn{\frac{1}{s}+B+C}
   = \frac{2\sinh^2\prn{\frac{m \beta}{2}}}{m s \cosh\prn{m \beta}}
   = \frac{2 S\prn{m \beta}}{M s \prn{S\prn{m \beta} + 1}}.
\end{equation}
%

>>>>

% \DeclareFontSubstitution{LS2}{stix}{m}{n}
% \DeclareSymbolFont{xlargesymbols}{LS2}{stixex}{m}{n}
% \DeclareMathDelimiter{\lAngle}{\mathopen}{xlargesymbols}{"EC}{xlargesymbols}{"12}
% \DeclareMathDelimiter{\rAngle}{\mathclose}{xlargesymbols}{"ED}{largesymbols}{"13}
% \DeclareMathDelimiter{\lVvert}{\mathopen}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
% \DeclareMathDelimiter{\rVvert}{\mathclose}{xlargesymbols}{"F2}{xlargesymbols}{"F5}
\usepackage{MnSymbol}
\newcommand{\aav}[1]{\mathopen{}\left\llangle #1 \right\rrangle\mathclose{}}
\newcommand{\nnrm}[1]{\mathopen{}\left\lWavy #1 \right\rWavy\mathclose{}}
%================================================================================
%
% \usepackage{adjustbox}
% \newcommand{\aligntop}[1]{\adjustbox{valign=t}{#1}}
% \newcommand{\alignmid}[1]{\adjustbox{valign=c}{#1}}
% %
% \usepackage{paralist}
% \newenvironment{myenuma}{\begin{inparaenum}[(a)]}{\end{inparaenum}}
% \newenvironment{myenumi}{\begin{inparaenum}[i)]}{\end{inparaenum}}
% \newenvironment{myenumA}{\begin{inparaenum}[\bfseries A.]}{\end{inparaenum}}
% \newenvironment{myenumI}{\begin{inparaenum}[\bfseries I.]}{\end{inparaenum}}
% \newenvironment{myenumr}{\begin{inparaenum}[1.]}{\end{inparaenum}}
% %for nested lists, \ref only produces innermost item:  (hacky)
% \makeatletter % for internal macros with @
% \renewcommand\p@enumii{}
% \makeatother
%
%================================================================================
%matrices
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MMm}{Q}
\newcommand{\MM}{\mathbf{\MMm}}
%probability distribution
\newcommand{\prm}{p}
\newcommand{\pr}{\mathbf{\prm}}
%equilibrium distribution
\newcommand{\eqm}{\pi}
\newcommand{\eq}{\boldsymbol{\eqm}}
%equilibrium metric
\newcommand{\Eqm}{\Pi}
\newcommand{\Eq}{\boldsymbol{\Eqm}}
%first passage times
\newcommand{\fptm}{T}
\newcommand{\fpt}{\mathbf{\fptm}}
%off-diagonal first passage times
\newcommand{\fptbm}{\overline{\fptm}}
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fundm}{Z}
\newcommand{\fund}{\mathbf{\fundm}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\arowm}{\xi}
\newcommand{\arow}{\boldsymbol{\arowm}}
\newcommand{\Lbm}{\Lambda}
\newcommand{\Lb}{\boldsymbol{\Lbm}}
\newcommand{\lbm}{\lambda}
\newcommand{\lb}{\boldsymbol{\lbm}}
\newcommand{\etwm}{\eta^w}
\newcommand{\etw}{\boldsymbol{\eta}^w}
\newcommand{\thbm}{\theta}
\newcommand{\thb}{\boldsymbol{\thbm}}
\newcommand{\ombm}{\omega}
\newcommand{\omb}{\boldsymbol{\ombm}}
\newcommand{\albm}{\alpha}
\newcommand{\alb}{\boldsymbol{\albm}}
\newcommand{\dgnm}{\zeta}
\newcommand{\dgn}{\boldsymbol{\dgnm}}
%synapse models
\newcommand{\wm}{w}
\newcommand{\w}{\mathbf{\wm}}
\newcommand{\Wm}{W}
\newcommand{\W}{\mathbf{\Wm}}
\newcommand{\MMdm}{M}
\newcommand{\MMd}{\mathbf{\MMdm}}
\newcommand{\encm}{K}
\newcommand{\enc}{\mathbf{\encm}}
\newcommand{\frgm}{\Wm^{\mathrm{F}}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\Fm}{\Phi}
\newcommand{\F}{\boldsymbol{\Fm}}
%eigenvectors
\newcommand{\evrm}{u}
\newcommand{\evr}{\mathbf{\evrm}}
\newcommand{\evlm}{\eta}
\newcommand{\evl}{\boldsymbol{\evlm}}
%KKT multipliers
\newcommand{\kktm}{\mu}
\newcommand{\kkt}{\boldsymbol{\kktm}}
%superscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
%snr curves etc
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{id}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\NNR}{NNR}
\DeclareMathOperator{\nnr}{NNR}
\newcommand{\snrb}{\overline{\snr}}
\DeclareMathOperator{\env}{Env}
\newcommand{\cnst}{\gamma}
% \newcommand{\CA}{\mathcal{A}}
% \newcommand{\CB}{\mathcal{B}}
% %avoid double superscripts
% \renewcommand{\pdiff}[2]{\frac{\partial #1}{\partial #2}}
% shifted process
\newcommand{\shift}[1]{\widehat{#1}}

@@

Note that
%
\begin{equation}\label{eq:laplaceeig}
  \hat{A}(s) = \sum_a \frac{\CI_a}{s + q_a},
\end{equation}
%
Therefore the eigenmodes with nonzero initial SNR can be found from singularity structure of \( \hat{A}(s) \) in the complex plane.
The poles are at \(s = - \frac{1}{\tau_a}\), with residue \(\CI_a\).

@@

So we see that all four inequalities cannot be simultaneously saturated, and the bound must be loose.
Because we think that \cref{eq:sqrtbound} can be improved, we will rewrite it as
%
\begin{equation}\label{eq:gammabnd}
  \CI_a \sqrt{r\tau_a} \leq \sqrt{\cnst}.
\end{equation}
% 
We have shown that the smallest \(\cnst\) for which \cref{eq:gammabnd} is true is less than 2.
Considering the two-state model with deterministic transitions, we see that \(\cnst\) is at least 1.
For the \(M\)-state uniform serial model with deterministic transitions, in \cref{eq:unieig}, we find
%
\begin{equation*}
  \CI_n \sqrt{r\tau_n} = 
      \frac{2 \sqrt{2} \sin\frac{n\pi}{2} \cos\frac{n\pi}{2M}}
           {\prn{M \sin \frac{n\pi}{2M}}^2},
\end{equation*}
%
which is maximised at \(n=1, M \to \infty\).
Thus we see that \(\cnst \geq \frac{128}{\pi^4} = 1.31 \).

@@

The first of these equations is solved by
%
\begin{equation}\label{eq:detagensol}
  \delta\etwm_i = B\e^{\beta(i-1)} + C\e^{\beta(1-i)},
  \qquad\text{where}\qquad
  s = S(\beta) \equiv 2 \sinh^2 \prn{\frac{\beta}{2}}
    = \cosh (\beta) - 1,
\end{equation}
%
with \(B\) and \(C\) determined by the boundary conditions provided by the last two equations of \eqref{eq:detahalf}:
%
\begin{equation}\label{eq:unicoeffs}
  B = -\frac{1}{s(1 + \e^{-\beta}) \cosh \prn{m \beta}},
  \qquad
  C = -\frac{1}{s(\e^{\beta} + 1) \cosh \prn{m \beta}}.
\end{equation}
%
This can be used to compute the Laplace transform
%
\begin{equation}\label{eq:unilaplace}
  A(s) = \frac{\etwm_1(s)}{m}
   = \frac{1}{m}\prn{\frac{1}{s}+B+C}
   = \frac{2\sinh^2\prn{\frac{m \beta}{2}}}{m s \cosh\prn{m \beta}}
   = \frac{2 S\prn{m \beta}}{M s \prn{S\prn{m \beta} + 1}}.
\end{equation}
%
This has poles as \( \beta = \frac{(2n-1)\pi\ir}{2m} \), therefore
%
\begin{equation}\label{eq:unieig}
  q_n = 1 - \cos\frac{(2n-1)\pi}{2m} = 2 \sin^2\frac{(2n-1)\pi}{4m},
  \qquad
  \CI_n = \frac{(-1)^{n-1} \sin\frac{(2n-1)\pi}{2m}}{m^2 q_n},
  \qquad n = 1,\dots,m.
\end{equation}
%
Note that these are only the eigenmodes with nonzero \(\CI_n\).
